### 2011-11-25 AffineLieAlgebras/Thesis/1.tex
More translations in chapter 1.

\subsubsection{Вложения и ортогональные пары подалгебр простых алгебр Ли}
\label{sect-embeddings}

In this subsection we discuss some properties of ''orthogonal pairs'' of
subalgebras in simple Lie algebras of classical series.

When both $\gf$ and $\af$ are finite-dimensional all regular
embeddings can be obtained by a successive elimination of nodes in the
extended Dynkin diagram of $\gf$ (and $\Delta_{\afb }^{+}=\emptyset $
if $\af$ is maximal). For the classical series $A$, $C$ and $D$ when
the regular injection $\af\rightarrow \gf$ is thus fixed, the
Dynkin diagram for $\afb$ is obtained from the extended diagram
of $\gf$ by eliminating the subdiagram of $\af$ and the adjacent
nodes:
\begin{table}[tbh]
\label{tab:diagrams} \noindent \centering{\
\begin{tabular}{|l|l|l|l|}
\hline
$\gf$ & Extended diagram of $\gf$ & Diagrams of the subalgebras $%
\af,\; \afb$ &  \\ \hline
$A_n$ & \includegraphics{table1_1(l)} & \includegraphics{table1_1(r)}
&  \\ \hline
$C_n$ & \includegraphics{table1_3(l)} & \includegraphics{table1_3(r)}
&  \\ \hline
$D_n$ & \includegraphics{table1_4(l)} & \includegraphics{table1_4(r)}
&  \\ \hline
\end{tabular}
}
\caption{Subalgebras $\af,\;\afb$ for the classical series}
\end{table}

In the case of $B$ series the situation is different. The reason is that
here the subalgebra $\afb$ may be larger than the one obtained
by elimination of the subdiagram of $\af$ and the adjacent nodes. The
subalgebras of the orthogonal pair,$\ \af$ and $\afb$, must
not form a direct sum in $\gf$ . It can be directly
checked that when $\gf=B_{r}$ and $\af=B_{r_{\af}}$ the
orthogonal subalgebra is $\afb=B_{r-r_{\af}}$. Consider the
injection $B_{r_{\af}}\rightarrow B_{r},\quad 1<r_{\af}<r$. By
eliminating the simple root $\alpha _{r_{\af}-1}=e_{r_{\af%
}-1}-e_{r_{\af}}$ one splits the extended Dynkin diagram of $B_{r}$
into the disjoint diagrams for $\af=B_{r_{\af}}$ and $D_{r-r_{%
\af}}$. But the system $\Delta _{\af_{\perp }}$ contains not only
the simple roots $\left\{ e_{1}-e_{2},e_{2}-e_{3},\ldots ,e_{r_{\af%
}-2}-e_{r_{\af}-1},e_{1}+e_{2}\right\} $ but also the root $e_{r_{\frak{%
a}}-1}$. Thus $\Delta _{\af_{\perp }}$ forms the subsystem of the type $%
B_{r-r_{\af}}$ and the orthogonal pair for the injection $B_{r_{\af%
}}\rightarrow B_{r}$ is $\left( B_{r_{\af}},B_{r-r_{\af}}\right) $%
. In the next Section a particular case of such orthogonal pair is
presented for the injection $B_{2}\rightarrow B_{4}$ (see Figure \ref
{fig:dynkin}).

The complete classification of regular subalgebras for affine Lie algebras
can be found in the recent paper \cite{1751-8121-41-36-365204}. From the complete
classification of maximal special subalgebras in classical Lie algebras \cite
{dynkin1952semisimple} we can deduce the following list of pairs of
orthogonal subalgebras $\af,\;\afb$:
\begin{equation*}
\begin{array}{lll}
su(p)\oplus su(q) & \subset su(pq) &  \\
so(p)\oplus so(q) & \subset so(pq) &  \\
sp(2p)\oplus sp(2q) & \subset so(4pq) &  \\
sp(2p)\oplus so(q) & \subset sp(2pq) &  \\
so(p)\oplus so(q) & \subset so(p+q) & \mathrm{{for}\;p\;{and}\;q\;{odd}.}
\end{array}
\end{equation*}

\subsubsection{Algorithm for recursive computation of branching coefficients}

@@

The recurrent relation (\ref{recurrent-relation}) allows us to formulate an
algorithm for recursive computation of branching coefficients. In this
algorithm there is no need to construct the module $L^{(\mu)}_{\gf}$ or
any of the modules $L^{(\nu)}_{\af}$.

It contains the following steps:

\begin{enumerate}
\item  Construct the root system $\Delta _{\af}$ for the embedding $%
\af\rightarrow \gf$.

\item  Select all positive roots $\alpha \in \Delta ^{+}$ orthogonal
to  $\af$, i.e. form the set $\Delta_{\afb }^{+}$.

\item  Construct the set $\Gamma _{\af\rightarrow \gf}$. Relation
 (\ref{eq:6}) defines the sign function
 $s(\gamma)$ and the set $\Phi_{\af\subset \gf}$ where the lowest weight
 $\gamma_0$ is to be subtracted to get the fan (\ref{fan-defined}):
 $\Gamma _{\af\rightarrow \gf}=\left\{ \xi -\gamma _{0}|\xi \in \Phi _{%
\af\subset \gf}\right\} \setminus \left\{ 0\right\}$.

\item  Construct the set $\widehat{\Psi ^{(\mu )}}=\left\{ w (\mu +\rho
)-\rho ;\;w \in W\right\} $ of singular weights for the $\gf$%
-module $L^{(\mu )}$.

\item  Select the weights $\left\{ \mu _{\widetilde{\af_{\perp }}%
}\left( w\right) =\pi _{\widetilde{\af_{\perp }}}\left[ w(\mu +\rho
)-\rho \right] -\mathcal{D}_{\af_{\perp }}\in \overline{C_{\widetilde{%
\af_{\perp }}}}\right\} $. Since the set $\Delta_{\afb }^{+}$ is fixed
we can easily check wether the weight $\mu _{\widetilde{\af_{\perp }}%
}\left( w\right) $ belongs to the main Weyl chamber $\overline{C_{\widetilde{%
\af_{\perp }}}}$ (by computing its scalar product with the fundamental
weights of $\afb^{+}$).

\item  For the weights $\mu _{\widetilde{\af_{\perp }}}\left( w\right) $
calculate dimensions of the corresponding modules, $\mathrm{\dim }\left(
L_{\widetilde{\af_{\perp }}}^{\mu _{\widetilde{\af_{\perp }}%
}\left( u\right) }\right) $, using the Weyl dimension formula and construct
the singular element $\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}$.

\item  Calculate the anomalous branching coefficients using the
recurrent relation (\ref{recurrent-relation}) and select among them those
corresponding to the weights in the main Weyl
chamber $\overline{C_{\af}}$.
\end{enumerate}

We can speed up the algorithm by
one-time computation of the representatives of the conjugate classes $%
W/W_{\afb }$.

The next section contains examples illustrating the application of this
algorithm.

>>>>

\subsubsection{Вложения и ортогональные пары подалгебр простых алгебр Ли}
\label{sect-embeddings}
В этом разделе мы обсудим некоторые свойства ``ортогональных пар'' подалгебр простых алгебр Ли классических серий.

Когда  $\gf$ и $\af$ конечномерные все регулярные вложения можно получить путем последовательного удаления вершин из расширенной диаграммы Дынкина алгебры $\gf$ (и  $\Delta_{\afb }^{+}=\emptyset $ если $\af$ максимальна). Если регулярное вложение для классических серий $A$, $C$ и $D$ зафиксировано таким образом, то диаграмма Дынкина для $\afb$ получается из расширенной диаграммы $\gf$ путем удаления поддиаграммы $\af$ и соседних вершин:
\begin{table}[tbh]
\label{tab:diagrams} \noindent \centering{\
\begin{tabular}{|l|l|l|l|}
\hline
$\gf$ & Расширенная диаграмма $\gf$ & Диаграммы подалгебр $\af,\; \afb$ &  \\ \hline
$A_n$ & \includegraphics{table1_1(l)} & \includegraphics{table1_1(r)}
&  \\ \hline
$C_n$ & \includegraphics{table1_3(l)} & \includegraphics{table1_3(r)}
&  \\ \hline
$D_n$ & \includegraphics{table1_4(l)} & \includegraphics{table1_4(r)}
&  \\ \hline
\end{tabular}
}
\caption{Подалгебры $\af,\;\afb$ для классических серий}
\end{table}

В случае серии $B$ ситуация отличается. Причина в том, что в данном случае подалгебра $\afb$ может быть больше, чем результат удаления поддиаграммы  $\af$ и соседних вершин. Подалгебры ортогональной пары $\ \af$ и $\afb$ не обязаны образовывать прямую сумму в $\gf$. Можно явно проверить, что при  $\gf=B_{r}$ и $\af=B_{r_{\af}}$ ортогональная подалгебра -- это $\afb=B_{r-r_{\af}}$. Рассмотрим вложение $B_{r_{\af}}\rightarrow B_{r},\quad 1<r_{\af}<r$. В результате удаления простого корня  $\alpha _{r_{\af}-1}=e_{r_{\af}-1}-e_{r_{\af}}$ расширенная диаграмма Дынкина для $B_{r}$ разбивается на несвязные диаграммы $\af=B_{r_{\af}}$ и $D_{r-r_{\af}}$. Однако корневая система  $\Delta _{\afb}$ содержит не только простые корни  $\left\{ e_{1}-e_{2},e_{2}-e_{3},\ldots ,e_{r_{\af}-2}-e_{r_{\af}-1},e_{1}+e_{2}\right\} $, но и корень $e_{r_{\af}-1}$. То есть  $\Delta _{\afb}$ образует систему типа  $B_{r-r_{\af}}$ и ортогональная пара для вложения $B_{r_{\af}}\rightarrow B_{r}$ -- это  $\left( B_{r_{\af}},B_{r-r_{\af}}\right) $. В следующем разделе мы приводим пример такой ортогональной пары для вложения $B_{2}\rightarrow B_{4}$ (см. Рис. \ref{fig:dynkin}).

Полная классификация регулярных подалгебр аффинных алгебр Ли приведена в недавней работе \cite{1751-8121-41-36-365204}. Из полной классификации максимальных специальных подалгебр классических алгебр Ли \cite{dynkin1952semisimple} мы получаем следующий список пар ортогональных подалгебр $\af,\;\afb$:
\begin{equation*}
\begin{array}{lll}
su(p)\oplus su(q) & \subset su(pq) &  \\
so(p)\oplus so(q) & \subset so(pq) &  \\
sp(2p)\oplus sp(2q) & \subset so(4pq) &  \\
sp(2p)\oplus so(q) & \subset sp(2pq) &  \\
so(p)\oplus so(q) & \subset so(p+q) & \text{для нечетных}\;p\;\text{и}\;q.
\end{array}
\end{equation*}

\subsubsection{Алгоритм рекурсивного вычисления коэффициентов ветвления}

@@

Рекуррентные соотношения (\ref{recurrent-relation}) позволяют нам сформулировать алгоритм для рекурсивного вычисления коэффициентов ветвления. При этом для работы алгоритма не требуется построение модуля $L^{(\mu)}_{\gf}$ или какого-то из модулей $L^{(\nu)}_{\af}$

Алгоритм состоит из следующих шагов:

\begin{enumerate}
\item Построить корневую систему $\Delta _{\af}$ для вложения $\af\rightarrow \gf$.

\item Выбрать все положительные корни $\alpha \in \Delta ^{+}$, ортогональные к  $\af$, то есть сформировать множество $\Delta_{\afb }^{+}$.

\item Построить множество $\Gamma _{\af\rightarrow \gf}$. Соотношение  (\ref{eq:6}) определяет знаковую функцию $s(\gamma)$ и множество $\Phi_{\af\subset \gf}$. Вычитание наименьшего веса $\gamma_0$ дает веер вложения (\ref{fan-defined}):
 $\Gamma _{\af\rightarrow \gf}=\left\{ \xi -\gamma _{0}|\xi \in \Phi _{\af\subset \gf}\right\} \setminus \left\{ 0\right\}$.

\item Построить множество $\widehat{\Psi ^{(\mu )}}=\left\{ w (\mu +\rho
)-\rho ;\;w \in W\right\} $ сингулярных весов для  $\gf$-модуля $L^{(\mu )}$.

\item Выбрать веса $\left\{ \mu _{\widetilde{\afb}}\left( w\right) =\pi _{\widetilde{\afb}}\left[ w(\mu +\rho
)-\rho \right] -\mathcal{D}_{\afb}\in \overline{C_{\widetilde{\afb}}}\right\} $. Так как множество  $\Delta_{\afb }^{+}$ фиксировано, легко осуществить проверку принадлежности веса $\mu _{\widetilde{\afb}}\left( w\right) $ к главной камере Вейля $\overline{C_{\widetilde{\afb}}}$ (достаточно вычислить скалярные произведения с фундаментальными весами $\afb^{+}$).

\item Для весов $\mu _{\widetilde{\afb}}\left( w\right) $ вычислить размерности соответствующих модулей, $\mathrm{\dim }\left(L_{\widetilde{\afb}}^{\mu _{\widetilde{\afb}}\left( u\right) }\right) $, используя формулу Вейля для размерностей. Затем построить сингулярный элемент $\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}$.

\item Вычислить аномальные коэффициенты ветвления используя рекуррентное соотношение (\ref{recurrent-relation}) и выбрать среди них те, которые соответствуют весам в главной камере Вейля $\overline{C_{\af}}$.
\end{enumerate}

Можно ускорить работу алгоритмы путем однократного вычисления представителей классов смежности $W/W_{\afb }$.

Следующий раздел состоит из примеров, иллюстрирующих применение данного алгоритма.
